Estimation of cloud-corrected microwave humidity radiances with associated uncertainties

### 2020-11-11 Article/ici_submm_cloudcorr.tex
included Dave's changes

Identification of cloud affected observations is indispensable for both clear-sky and all-sky numerical weather prediction (NWP) systems. In this study, we present a methodology to identify and correct the cloud impact in microwave humidity radiances. A channel specific quantile regression neural network (QRNN) based approach is formulated to estimate the posterior distributions of noise free clear-sky (NFCS) radiances. The applicability of the approach is demonstrated by application to a present sensor: MicroWave Humidity Sounder-2 (MWHS-2). With the future outlook, the applicability to sub-millimeter (sub-mm) channels is also analysed. The results are compared against the NFCS simulations. With the present sensors, the approach is partially successful in correcting the negative departures. Almost all cases with high cloud impact are corrected, but thin cirrus clouds cannot be fully corrected. In spite of the partial performance, the results from QRNN are better than what operational cloud filtering techniques like scattering index can achieve. On the other hand, when sub-mm observations from ICI are used to correct the cloud impact, QRNN can successfully correct most of the cases with cloud impact. Only 2-6\% of the cases are partially corrected and these can be easily removed by a simple filtering. The resulting error distributions are almost symmetric with a very low spread. The methodology also works well when only one of the higher frequencies (325\,GHz) is available. Except for the lowest peaking channel, using sub-mm observations also results in error distributions with standard deviation less than the measurement noise. Another advantage of using QRNN is that predicted quantiles can be employed to quantify the case-specific uncertainty estimates. The results show that QRNN is successful in representing the uncertainty for each case individually. Despite individual variations for each channel, the predictions exhibit a higher uncertainty when cloudy cases lack orthogonal information from other channels. In comparison to deterministic correction/filtering approaches, the high quality corrected radiances with uncertainty estimates have great potential to be used efficiently in assimilation systems.

@@

Another motivation necessitating the need for cloud clearing is usage of 183\,GHz channels in 
numerical weather prediction (NWP). Usage of passive microwave data by
``all-sky'' assimilation in global NWP is growing \citep{geer2017growing},
but 183\,GHz data are still mainly used in a clear-sky fashion
\citep{geer2018all}. The latter is particularly true in NWP of regional scope
\citep{gustafsson2018survey}, all-sky assimilation of 183\,GHz at many national
weather agencies will likely not be reached in many years. Anyhow, both clear-sky and all-sky assimilation require identification of cloud affected observations. ``Scattering index'' \citep{geer2015scatteringindex} is one of the approaches used. It is based on brightness temperature differences between 89 and 150\,GHz. Another commonly used cloud filtering method is based on ``observation minus background'' (O-B), where the forecast model is used to obtain an estimate of the expected clear-sky value and the observation is rejected if the deviation exceeds some threshold \citep{English1999clouddetection}.

Despite their broad usage, these filtering methods aim at partial removal of cloud contamination originating from high ice clouds. The residual cloud impact can have some drawbacks. For 183\,GHz, the impact of hydrometeors systematically causes a decrease in the observed radiance (maybe with exceptions at extremely dry conditions), see
e.g.\ \citet{barlakas:three:20} and below. This implies that if any cloud contamination is missed by the filtering this will cause a negative bias in mean radiance, compared to the true clear-sky mean. This bias will translate to a bias in humidity after the retrieval or assimilation. For NWP systems assimilating clear-sky observations, the effect of undetected clouds is overcome by inflating the observational errors. However this artificial suppression of observations diminishes the impact of observations. The alternative is to apply a very strict filtering, but this increases the rejection of clear-sky values, i.e.\ an important loss of useful data. Another limitation of existing clearing approaches is their ``one for all'' approach, i.e.\ observations in all 183\,GHz channels are either kept or rejected. This often rejects more observations than needed, as the channels differ in their altitude coverage. An observation could be cloudy in some channels and still the be clear-sky in others. To allow a channel specific filtering, data likely need to be combined in a more complex manner than simple differences, but it is unclear what type of regression that would be best. This points towards applying machine learning, as used by e.g.\ \citet{favrichon2019detecting}. A maybe less obvious problem is the assignment of uncertainty to the filtered values. To our best knowledge, so far only estimates of mean and worst case errors have been provided. Some cases with relatively high cloud impact will likely be missed, while most cases are clear-sky from start. As the remaining cloudy cases can cause significant biases, the likely solution is to apply a quite conservative (high) error estimate. However, this will unnecessary downgrade the value of the truly clear-sky cases and the observations are used in a non-optimal manner.

We are here approaching cloud filtering task from a new angle. The basic idea
is to derive an estimate of the corresponding noise-free clear-sky (NFCS) value
(i.e.\ the radiance that would have been measured in absence of noise and
hydrometeors). This is done for each channel separately, only using
measurements (no ``background'' data involved). Not only a best estimate is
provided, but also a case specific uncertainty. This information could be used as a pure filter, by rejecting data where the correction exceeds some threshold value. However, even
better is to replace the original value with the predicted NFCS value when
forming the clear-sky dataset. We denote this approach as cloud correction. It
is shown below that a basically bias free cloud correction can be obtained.
This feature also removes the need of threshold value, as long as the retrieval or
assimilation system can incorporate the uncertainty of the corrected value. As
also will be shown, the uncertainty for originally clear-sky data is determined
by noise, but the uncertainty increases with magnitude of correction.
Accordingly, the cloud correction approach allows that the full weight of
clear-sky data can be preserved.

@@

The main focus of this study is the potential of proposed cloud-correction
method using sub-millimetre (sub-mm) observations, which will become available
operationally with the launch of the Ice Cloud Imager (ICI,
\citet{eriksson:towar:20}) on board the next generation of EUMETSAT Polar System - Second Generation (EPS-SG), but will likely also be included in
several smaller missions (such as Artic Weather Satellite (AWS); presented
below). Additionally, we demonstrate the feasibility of the approach based on
observations at 89 and 150\ GHz channels (following
\citet{geer2015scatteringindex}), which are available already today. The focus
on sub-mm channels is motivated by several reasons. Firstly, the higher
frequencies are more sensitive to scattering effects from smaller
hydrometeors. For high altitude cirrus clouds, these frequencies have a stronger
signal than 183\,GHz. In some cloudy situations it could be possible that the
impact at 183\,GHz is of the order of thermal noise and modelling uncertainties,
while the impact at 325\,GHz is significantly higher. Secondly, the proposed
cloud correction methods allows integrating ICI sub-mm observations in
clear-sky data assimilation schemes and thus provides a simple
way to make use this novel data source as soon as it becomes available.

@@

\section{Data and methods}
\label{data_methods}
%
\subsection{Satellite Instruments}
%
\subsubsection{ MicroWave Humidity Sounder-2}
%
%t
\begin{table}[t]
	\caption{Specifications of MWHS-2 channels relevant to this study.}
	\label{tab:specifications_MWHS2}	
	\begin{tabular}{crrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
		& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		1	&	89.0   		  & 1500			&	1.0	\\
		6	&	118.75±1.1    & \phantom{0}200 	&	1.6\\
		7	&	118.75±2.5    & \phantom{0}200 	&	1.6\\
		10	&	150.0         & 1500 			&	1.0 \\
		11	&	183.31±1      & \phantom{0}500  &	1.0 \\
		12  & 	183.31±1.8    & \phantom{0}700 	&   1.0\\
		13  & 	183.31±3      & 1000    		&	1.0	\\
		14  & 	183.31±4.5    & 2000    		&	1.0\\
		15  & 	183.31±7      & 2000  			&	1.0  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
MicroWave Humidity Sounder 2 (MWHS-2) is an instrument onboard FengYun-3C (FY-3C). MWHS-2 a cross track scanning microwave radiometer and measures 15 frequencies in the range 89-191\,GHz. 89\,GHz and 150\,GHz are window channels, five humidity sounding channels are centered at 183.31\,GHz, and eight temperature sounding channels are centered at 118.75\,GHz oxygen absorption line. The five humidity sounding channels are similar to ATMS. As of December 2019, observations from MWHS-2 are routinely assimilated at European Centre for Medium-Range Weather Forecasts (ECMWF, \citet{duncan2020MWHS}). The channels relevant to this study are described on Table~\ref{tab:specifications_MWHS2}.

For this study, MWHS-2 simulations were sourced from ECMWF assimilation system. \todo{details?}

@@

\subsection{Simulations}
%
MWHS-2 simulated radiances are routinely calculated at ECMWF for their use in the ECMWF NWP system. RTTOV (Radiative Transfer for TOVS, \citet{saunders2018update} and RTTOV-SCATT \citep{bauer2006rttovscat} are used to calculate the radiances in clear-sky and all-sky conditions, respectively. For this study, MWHS-2 brightness temperatures for the period June-July 2020 are used. Out of all the available observations, we use data for latitudinal range $60^{\degree}$\,S to $60^{\degree}$\,N and satellite zenith angle less than $7.5^{\degree}$. With this filter we have 340\,000 cases.

@@

All evaluation results except hyper-parameter tuning, described in the study, have been made on the test dataset. Hyper-parameter tuning is made on the training dataset (see Appendix~\ref{appendix:hyperparamter}).

\section{Correcting cloud affected data in MWHS-2}
\label{qrnn_mwhs}
%
In this section, we introduce the QRNN based cloud correction in the context of current operational sensors.  We use MWHS-2 to demonstrate the results. The choice is motivated by the fact that MWHS-2 has five complementary 183\,GHz channels, along with additional 118\,GHz channels. In order to formulate and test the correction approach, multiple QRNN experiments are performed for MWHS-2 channel 14. Later the optimal experiment is extended to other 183\,GHz channels. A brief comparison of the results is also made against existing cloud filtering methods. Further, the estimates of case-specific uncertainties obtained from QRNN are also evaluated.

@@

	\item In the first experiment, we examine the performance of QRNN cloud correction with MWHS-2 window channels: 89 and 150\,GHz. In ECMWF NWP system, the differences between the observations of these two window channels are used to identify the cloud affected data for ATMS and MWHS-2 \citep{geer2015scatteringindex}. To investigate the potential impact of these two window channels in QRNN based cloud correction, the configuration QRNN-single is applied. In this experiment, the training inputs include all-sky brightness temperatures from target 183\,GHz channel, 89\,GHz and 150\,GHz. Both 89 and 150\,GHz are window channels so the land sea mask is also included as a training input. For example, for channel 14, the training inputs are all-sky brightness temperatures from channels 14, 1, 10 and land/sea mask. This combination is referred as 89+150\,GHz in the text.

@@

\subsubsection{QRNN-single applied to MWHS-2 channel 14} 
\begin{figure}[t]
	\centering
	\includegraphics[width = 70mm]{Figures/QRNN_output_mwhs.pdf} 
	\caption{Distribution of point estimate $\ynfcs$ obtained from QRNN-single 89+150\,GHz for MWHS-2 channel 14. The corresponding distributions for all-sky and clear-sky simulations are also shown.}
	\label{fig:distribution_predicted_mwhs14}	
\end{figure}
\begin{figure}[t]
	\centering
	\includegraphics[width = 70mm]{Figures/MWHS_error_dist_14.pdf} 
	\caption{Error distribution for deviations of point estimates to NFCS simulations. The results are from QRNN-single experiment 89+150\,GHz and MWHS-2 channel 14. Noise is also plotted for reference. The label ``All-sky'' represents the all-sky simulations, ``Predicted(All)'' denotes the predicted point estimates. The error distribution achieved with scattering index (SI) filtering is also shown. }
	\label{fig:error_distribution_mwhs14}	
\end{figure}
\begin{table*}[t]
	\caption{Error statistics for deviations of point estimates to NFCS simulations. Results are for different QRNN experiments for MWHS-2 channel 14 (see Sect.~\ref{sec:QRNN_expt_MWHS}). The statistics for all-sky and clear-sky simulations are also provided. The label ``All'' denotes the entire dataset of predicted point estimates, while ``Pred.(5\,K)'' refers to the predicted point estimates but where cases with cloud correction greater than 5\,K are excluded. The last two columns show the statistics obatined after filtering cloudy cases according to scattering index (SI) and scheme by \citet{buehler:aclou:07} (B183). Bias, MAE, STD are in K, skewness in dimensionless.}
	\label{tab:error_statistics_mwhs_14}
	\setlength{\tabcolsep}{4pt}
	\begin{tabular}{lrr|rr|rr|rr|rr|rr}
		\tophline
		&\multicolumn{2}{c|}{Simulations}& \multicolumn{6}{c|}{QRNN-single} & \multicolumn{2}{c|}{QRNN-all} & \multicolumn{2}{c}{Pure filtering}\\
		\cline{2-13}
		%		\hline
		&   Clear-sky &   All-sky &  \multicolumn{2}{c|}{89+150\,GHz} & \multicolumn{2}{c|}{89+150+118\,GHz} & \multicolumn{2}{c|}{150\,GHz} & \multicolumn{2}{c|}{89+150+183\,GHz}& SI & B183\\		
		&			   &			& All & Pred.(5\,K) & All & Pred.(5\,K) & All & Pred.(5\,K)  & All & Pred.(5\,K)&&\\
		\middlehline
Bias     & -0.00 &  -0.83 & -0.09 & -0.08 & -0.04 & -0.03 & -0.09 & -0.09 & -0.09 & -0.08 & -0.24 & -0.51\\
MAE      &  0.80 &   1.44 &  0.88 &  0.86 &  0.89 &  0.87 &  0.89 &  0.87 &  0.62 &  0.59 &  0.92 &  1.14\\
STD      &  1.00 &   3.89 &  1.19 &  1.14 &  1.19 &  1.15 &  1.21 &  1.16 &  0.90 &  0.83 &  1.26 &  1.83\\
Skewness &  0.01 & -14.40 & -1.11 & -0.96 & -1.02 & -0.86 & -0.89 & -1.02 & -1.61 & -1.58 & -1.79 & -3.45\\
Rejection&  -	 & - 	  & - 	  & 3.3\% & - 	  & 3.4\% & -     & 3.2\% & -     & 3.3\% & 28.1\%& 3.5\%\\
\bottomhline
	\end{tabular}
\end{table*}

@@

\subsubsection{QRNN-single applied to channel 11, 12, 13 and 15}
\label{sec:mwhs_others}
\begin{table*}[t]
	\caption{ Error statisics for MWHS-2 channels 11, 12, 13 and 15 from QRNN-single experiment 89+150\,GHz. The labels ``All'' and ``Pred.(5\,K)' are same as described in Table~\ref{tab:error_statistics_mwhs_14}. The statistics from existing filtering approaches: scattering index (SI), and scheme by \citet{buehler:aclou:07} (B183) are also provided.}
	\label{tab:error_statistics_mwhs_others}
	\begin{tabular}{lrrr|rr|rr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& \multicolumn{2}{c|}{QRNN-single} & \multicolumn{2}{c}{Pure filtering} \\
		\cline{2-8}
		%		\hline
		& &  Clear-sky &   All-sky &  \multicolumn{2}{c|}{89+150\,GHz}    & SI & B183  \\
		&	&		   &			& All & Pred.(5K) &&\\
		\middlehline
		Channel 11  &   Bias      & 0.00 &  -0.16 & -0.03 & -0.03		 &-0.04	& -0.05 \\
				    &	MAE       & 0.80 &   0.89 &  0.79 &  0.79  		 & 0.81	&  0.81 \\
					&	STD       & 1.00 &   1.71 &  1.03 &  1.01 		 & 1.02	&  1.03 \\
					&	Skewness  & 0.01 & -20.18 & -0.68 & -0.45 		 &-0.12 & -0.24 \\
					& Rejection   & -    & -      & -	  & 0.2\%        & 28.1\% & 3.5\%  \\
		\middlehline
		Channel 12  & Bias      & -0.00 &  -0.28 & -0.05 & -0.05		&  -0.07 & -0.15 \\
					& MAE       &  0.80 &   0.98 &  0.81 &  0.80  		&   0.82 &	0.86 \\
					& STD       &  1.00 &   2.26 &  1.07 &  1.05  		&  1.04  &  1.14 \\
					& Skewness  &  0.00 & -20.72 & -0.91 & -0.70  		&  -0.38 & -0.86 \\
					& Rejection & -    & -      & -	  & 0.6\%        & 28.1\% & 3.5\%  \\
		\middlehline
		Channel 13  & Bias      & -0.00 &  -0.52 & -0.08 & -0.07 		& -0.14  & -0.29 \\
					& MAE       &  0.80 &   1.17 &  0.84 &  0.83  		&  0.86  &  0.97 \\
					& STD       &  1.00 &   3.04 &  1.13 &  1.10  		&  112  &	1.39 \\
					& Skewness  &  0.02 & -17.85 & -1.14 & -0.98  		& -0.91  & -2.23 \\	
					& Rejection   & -    & -      & -	  & 1.6\%        & 28.1\% & 3.5\%  \\	
		\middlehline
		Channel 15  & Bias      & -0.00 &  -1.28 & -0.07 & -0.06        & -0.33 &  -0.81 \\
					& MAE       &  0.80 &   1.87 &  0.97 &  0.92 		&  1.03 &	1.44 \\
					& STD       &  1.00 &   5.08 &  1.35 &  1.25 		&  1.53 &	2.64 \\
					& Skewness  & -0.00 & -11.15 & -0.60 & -0.83 		& -2.85 &  -4.41 \\ 
					& Rejection & -     & -      & -	 & 5.6\%        & 28.1\% & 3.5\%  \\
		\bottomhline
	\end{tabular}
\end{table*}

@@

The symmetric and low spread error distribution with uncertainty estimates is an important result from the DA perspective. Most of the existing cloud filtering schemes work well only at removing cases with high cloud impact, and as a consequence, the error distributions are highly skewed. To use these observations correctly, data assimilation schemes often inflate their asymmetric error distributions at the cost of artificially suppressing the observational impact. This is not the optimal solution but the best which can be achieved with present sensors. However, the symmetric error distributions obtained from QRNN allow us to effectively utilize more than 94\% of the data without the need for error inflation.

>>>>

A methodology based on quantile regression neural networks (QRNN) is presented that identifies and corrects the cloud impact on microwave humidity sounder radiances at 183GHz. This approach estimates the posterior distributions of noise free clear-sky (NFCS) radiances, providing nearly bias-free estimates of clear-sky radiances with a full posterior error distribution. It is first demonstrated by application to a present sensor, the MicroWave Humidity Sounder-2 (MWHS-2), then the applicability to sub-millimeter (sub-mm) sensors is also analysed. The QRNN results improve upon what operational cloud filtering techniques like a scattering index can achieve, but are ultimately imperfect due to limited information content on cirrus impact from traditional microwave channels---the negative departures associated with high cloud impact are successfully corrected, but thin cirrus clouds cannot be fully corrected. In contrast, when sub-mm observations are used, QRNN successfully corrects most cases with cloud impact, with only 2-6\% of cases left partially corrected. The methodology works well even if only one sub-mm channel (325\,GHz) is available. When using sub-mm observations, cloud correction usually results in error distributions with standard deviation less than typical channel noise values. Furthermore, QRNN outputs predicted quantiles for case-specific uncertainty estimates, successfully representing the uncertainty of cloud correction for each observation individually. In comparison to deterministic correction or filtering approaches, the corrected radiances and attendant uncertainty estimates have great potential to be used efficiently in assimilation systems due to being largely unbiased and adding little further uncertainty to the measurements.

@@

Another motivation necessitating the need for cloud clearing is usage of 183\,GHz channels in 
numerical weather prediction (NWP). Usage of passive microwave data by
``all-sky'' assimilation in global NWP is growing \citep{geer2017growing},
but 183\,GHz data are still mainly used in a clear-sky fashion
\citep{geer2018all}. The latter is particularly true in NWP of regional scope
\citep{gustafsson2018survey}, with clear-sky assimilation of 183GHz radiances still commonplace. Regardless, both clear-sky and all-sky assimilation require identification of cloud affected observations, either to screen out these observations or to assign an appropriate observation error. ``Scattering index'' \citep{geer2015scatteringindex} is one of the approaches used. It is based on brightness temperature differences between 89 and 150\,GHz. Another commonly used cloud filtering method is based on ``observation minus background'' (O-B), where the forecast model is used to obtain an estimate of the expected clear-sky value and the observation is rejected if the deviation exceeds some threshold \citep{English1999clouddetection}.

%Despite their broad usage, these filtering methods aim at partial removal of cloud contamination originating from high ice clouds. The residual cloud impact can have some drawbacks. 
At 183\,GHz, the impact of hydrometeors typically causes a decrease in the observed radiance due to scattering from ice hydrometeors (e.g.\ \citet{barlakas:three:20}). This implies that if any cloud contamination is missed by the filtering, a negative bias in mean radiance will result compared to the true clear-sky mean, which may translate into a bias in humidity after the retrieval or assimilation. For NWP systems assimilating clear-sky observations, the effect of undetected clouds may be overcome by inflating the observational errors and diminishing the impact of observations. Furthermore, the mathematical assumptions of data assimilation are predicated on Gaussian errors with no mean bias, and residual cloud impacts that cause a net bias are not easily handled by variational bias correction. One solution is to apply a very strict filtering, but this increases the rejection of clear-sky values, i.e.\ an important loss of useful data. Another limitation of existing clearing approaches is their ``one for all'' approach, i.e.\ observations in all 183\,GHz channels are either kept or rejected. This often rejects more observations than needed, as the channels differ in their altitude coverage. An observation could be cloudy in some channels and still the be clear-sky in others. To allow a channel specific filtering, data likely need to be combined in a more complex manner than simple differences, but it is unclear what type of regression would be best as the ideal solution would be scene-dependent. This points towards applying machine learning, as used by e.g.\ \citet{favrichon2019detecting}. A maybe less obvious problem is the assignment of uncertainty to the filtered values. To our best knowledge, so far only estimates of mean and worst case errors exist in the literature. Some cases with relatively high cloud impact will likely be missed, while most cases are clear-sky from the start. As the remaining cloudy cases can cause significant biases, the likely solution is to apply a quite conservative (high) error estimate. However, this will unnecessarily downgrade the value of the truly clear-sky cases and the observations are used in a non-optimal manner.

In this study, we approach the cloud filtering task from a new angle. The basic idea
is to derive an estimate of the corresponding noise-free clear-sky (NFCS) value
(i.e.\ the radiance that would have been measured in absence of noise and
hydrometeors). This is done for each channel separately, only using
measurements (no ``background'' data involved). Not only a best estimate is
provided, but also a case-specific uncertainty. This information could be used as a pure filter, by rejecting data where the correction exceeds some threshold value. However, even
better is to replace the original value with the predicted NFCS value when
forming the clear-sky dataset. We denote this approach as cloud correction. It
is shown below that a basically bias free cloud correction can be obtained.
This feature also removes the need for defining threshold values, as long as the retrieval or
assimilation system can incorporate the uncertainty of the corrected value. As
also will be shown, the uncertainty for originally clear-sky data is determined
by noise, but the uncertainty increases with magnitude of correction.
Accordingly, the cloud correction approach permits the full weight of
clear-sky data to be preserved.

@@

The main focus of this study is the potential of this cloud-correction
method using sub-millimetre (sub-mm) observations, which will become available
operationally with the launch of the Ice Cloud Imager (ICI,
\citet{eriksson:towar:20}) on board the next generation of EUMETSAT Polar System - Second Generation (EPS-SG), but will likely also be included in
several smaller missions (such as Artic Weather Satellite (AWS); presented
below). Additionally, we demonstrate the feasibility of the approach based on
observations at 89 and 150\ GHz channels (following
\citet{geer2015scatteringindex}), which are available on several sensors extant today. The focus
on sub-mm channels is motivated by several reasons. First, the higher
frequencies are more sensitive to scattering effects from smaller
hydrometeors and are thus expected to provide greater sensitivity to high altitude cirrus clouds. For example, in some cloudy situations the
cloud impact at 183\,GHz may be of the order of thermal noise and modelling uncertainties,
while the impact at 325\,GHz is significant enough to provide sufficient signal to noise for identifying cloud. Second, the proposed
cloud correction methods allow integration of ICI sub-mm observations in
clear-sky data assimilation schemes with no further modifications, thus providing a simple
way to make use this novel data source as soon as it becomes available.

@@

\section{Data and methods}
\label{data_methods}
%
\subsection{Satellite Instruments}
%
\subsubsection{ MicroWave Humidity Sounder-2}
%
%t
\begin{table}[t]
	\caption{Specifications of MWHS-2 channels relevant to this study.}
	\label{tab:specifications_MWHS2}	
	\begin{tabular}{crrr}
		\tophline
		Channel & Frequency 	& Bandwidth & NE$\Delta$T \\
		& [GHz]			& [MHz]		& [K]		\\
		\middlehline
		1	&	89.0   		  & 1500			&	1.0	\\
		6	&	118.75±1.1    & \phantom{0}200 	&	1.6\\
		7	&	118.75±2.5    & \phantom{0}200 	&	1.6\\
		10	&	150.0         & 1500 			&	1.0 \\
		11	&	183.31±1      & \phantom{0}500  &	1.0 \\
		12  & 	183.31±1.8    & \phantom{0}700 	&   1.0\\
		13  & 	183.31±3      & 1000    		&	1.0	\\
		14  & 	183.31±4.5    & 2000    		&	1.0\\
		15  & 	183.31±7      & 2000  			&	1.0  \\
		\bottomhline
	\end{tabular}
	\belowtable{} % Table Footnotes
\end{table}
The MicroWave Humidity Sounder 2 (MWHS-2) is an instrument on two current satellites in the FengYun-3 series, FY-3C and FY-3D. MWHS-2 a cross track scanning microwave radiometer and measures 15 frequencies in the range 89-191\,GHz. 89\,GHz and 150\,GHz are window channels, five humidity sounding channels are centered at 183.31\,GHz, and eight temperature sounding channels are centered on the 118.75\,GHz oxygen absorption line. The five humidity sounding channels are similar to ATMS. Observations from MWHS-2 are routinely assimilated in all-sky conditions at the European Centre for Medium-Range Weather Forecasts (ECMWF) with demonstrable positive impact on forecast performance \citep{duncan2020MWHS}. The channels relevant to this study are described in Table~\ref{tab:specifications_MWHS2}.

For this study, MWHS-2 simulations were sourced from the operational ECMWF assimilation system.

@@

\subsection{Simulations}
%
MWHS-2 simulated radiances during the period June-July 2020, are sourced from ECMWF. In the current version of the ECMWF Integrated Forecasting System (IFS), cycle 47R1 \citep{IFS47R1chap1}, clear-sky and all-sky radiative transfer are performed simultaneously for monitoring purposes, despite all humidity sounders being assimilated via all-sky exclusively. These side by side radiative transfer calculations on a large variety of model scenes from the IFS provides an ideal dataset for comparing radiances with and without cloud effects. Out of all the available observations during the period, we use data for latitudinal range $60^{\degree}$\,S to $60^{\degree}$\,N and satellite zenith angle less than $7.5^{\degree}$. With this filter we have approximately 290\,000 cases.

@@

All evaluation results except hyper-parameter tuning, described in the study, have been made on the test dataset. Hyper-parameter tuning is made on the validation dataset (see Appendix~\ref{appendix:hyperparamter}). Validation dataset is part of the training dataset which is held back during the training process.

\section{Correcting cloud affected data in MWHS-2}
\label{qrnn_mwhs}
%
In this section, we introduce the QRNN based cloud correction in the context of current operational sensors.  We use MWHS-2 to demonstrate the results. The choice is motivated by the fact that MWHS-2 has five complementary 183\,GHz channels, along with additional 118\,GHz channels. In order to formulate and test the correction approach, multiple QRNN experiments are performed for MWHS-2. However, for brevity we compare the comprehensive results only for channel 14. Later the optimal experiment is extended to other 183\,GHz channels. A brief comparison of the results is also made against existing cloud filtering methods. Further, the estimates of case-specific uncertainties obtained from QRNN are also evaluated.

@@

	\item In the first experiment, we examine the performance of QRNN cloud correction with MWHS-2 window channels: 89 and 150\,GHz. In ECMWF NWP system, the differences between the observations of these two window channels are used to identify the cloud affected data for humidity sounding channels \citep{geer2015scatteringindex}. To investigate the potential impact of these two window channels in QRNN based cloud correction, the configuration QRNN-single is applied. In this experiment, the training inputs include all-sky brightness temperatures from target 183\,GHz channel, 89\,GHz and 150\,GHz. Both 89 and 150\,GHz are window channels so the land sea mask is also included as a training input. For example, for channel 14, the training inputs are all-sky brightness temperatures from channels 14, 1, 10 and land/sea mask. This combination is referred as 89+150\,GHz in the text.

@@

\subsubsection{QRNN-single applied to MWHS-2 channel 14} 
\begin{figure}[t]
	\centering
	\includegraphics[width = 70mm]{Figures/QRNN_output_mwhs.pdf} 
	\caption{The distribution of point estimate $\ynfcs$ obtained from QRNN-single 89+150\,GHz for MWHS-2 channel 14. The corresponding distributions for all-sky and clear-sky simulations are also shown.}
	\label{fig:distribution_predicted_mwhs14}	
\end{figure}
\begin{figure}[t]
	\centering
	\includegraphics[width = 70mm]{Figures/MWHS_error_dist_14.pdf} 
	\caption{The error distribution for deviations of point estimates to NFCS simulations. The results are from QRNN-single experiment 89+150\,GHz and MWHS-2 channel 14. Noise is also plotted for reference. The label ``All-sky'' represents the all-sky simulations, ``Predicted(All)'' denotes the predicted point estimates. The error distribution achieved with scattering index (SI) filtering is also shown. }
	\label{fig:error_distribution_mwhs14}	
\end{figure}
\begin{table*}[t]
	\caption{The error statistics for deviations of point estimates to NFCS simulations. Results are for different QRNN experiments for MWHS-2 channel 14 (see Sect.~\ref{sec:QRNN_expt_MWHS}). The statistics for all-sky and clear-sky simulations are also provided. The label ``All'' denotes the entire dataset of predicted point estimates, while ``Pred.(5\,K)'' refers to the predicted point estimates but where cases with cloud correction greater than 5\,K are excluded. The last two columns show the statistics obtained after filtering cloudy cases according to scattering index (SI) and scheme by \citet{buehler:aclou:07} (B183). Bias, MAE, STD are in K, skewness in dimensionless.}
	\label{tab:error_statistics_mwhs_14}
	\setlength{\tabcolsep}{4pt}
	\begin{tabular}{lrr|rr|rr|rr|rr|rr}
		\tophline
		&\multicolumn{2}{c|}{Simulations}& \multicolumn{6}{c|}{QRNN-single} & \multicolumn{2}{c|}{QRNN-all} & \multicolumn{2}{c}{Pure filtering}\\
		\cline{2-13}
		%		\hline
		&   Clear-sky &   All-sky &  \multicolumn{2}{c|}{89+150\,GHz} & \multicolumn{2}{c|}{89+150+118\,GHz} & \multicolumn{2}{c|}{150\,GHz} & \multicolumn{2}{c|}{89+150+183\,GHz}& SI & B183\\		
		&			   &			& All & Pred.(5\,K) & All & Pred.(5\,K) & All & Pred.(5\,K)  & All & Pred.(5\,K)&&\\
		\middlehline
Bias     & -0.00 &  -0.83 & -0.09 & -0.08 & -0.04 & -0.03 & -0.09 & -0.09 & -0.09 & -0.08 & -0.24 & -0.51\\
MAE      &  0.80 &   1.44 &  0.88 &  0.86 &  0.89 &  0.87 &  0.89 &  0.87 &  0.62 &  0.59 &  0.92 &  1.14\\
STD      &  1.00 &   3.89 &  1.19 &  1.14 &  1.19 &  1.15 &  1.21 &  1.16 &  0.90 &  0.83 &  1.26 &  1.83\\
Skewness &  0.01 & -14.40 & -1.11 & -0.96 & -1.02 & -0.86 & -0.89 & -1.02 & -1.61 & -1.58 & -1.79 & -3.45\\
Rejection&  -	 & - 	  & - 	  & 3.3\% & - 	  & 3.4\% & -     & 3.2\% & -     & 3.3\% & 28.1\%& 3.5\%\\
\bottomhline
	\end{tabular}
\end{table*}

@@

\subsubsection{QRNN-single applied to channel 11, 12, 13 and 15}
\label{sec:mwhs_others}
\begin{table*}[t]
	\caption{ Error statistics for MWHS-2 channels 11, 12, 13 and 15 from QRNN-single experiment 89+150\,GHz. The labels ``All'' and ``Pred.(5\,K)' are same as described in Table~\ref{tab:error_statistics_mwhs_14}. The statistics from existing filtering approaches: scattering index (SI), and scheme by \citet{buehler:aclou:07} (B183) are also provided.}
	\label{tab:error_statistics_mwhs_others}
	\begin{tabular}{lrrr|rr|rr}
		\tophline
		&&\multicolumn{2}{c|}{Simulations}& \multicolumn{2}{c|}{QRNN-single} & \multicolumn{2}{c}{Pure filtering} \\
		\cline{2-8}
		%		\hline
		& &  Clear-sky &   All-sky &  \multicolumn{2}{c|}{89+150\,GHz}    & SI & B183  \\
		&	&		   &			& All & Pred.(5\,K) &&\\
		\middlehline
		Channel 11  &   Bias      & 0.00 &  -0.16 & -0.03 & -0.03		 &-0.04	& -0.05 \\
				    &	MAE       & 0.80 &   0.89 &  0.79 &  0.79  		 & 0.81	&  0.81 \\
					&	STD       & 1.00 &   1.71 &  1.03 &  1.01 		 & 1.02	&  1.03 \\
					&	Skewness  & 0.01 & -20.18 & -0.68 & -0.45 		 &-0.12 & -0.24 \\
					& Rejection   & -    & -      & -	  & 0.2\%        & 28.1\% & 3.5\%  \\
		\middlehline
		Channel 12  & Bias      & -0.00 &  -0.28 & -0.05 & -0.05		&  -0.07 & -0.15 \\
					& MAE       &  0.80 &   0.98 &  0.81 &  0.80  		&   0.82 &	0.86 \\
					& STD       &  1.00 &   2.26 &  1.07 &  1.05  		&  1.04  &  1.14 \\
					& Skewness  &  0.00 & -20.72 & -0.91 & -0.70  		&  -0.38 & -0.86 \\
					& Rejection & -    & -      & -	  & 0.6\%        & 28.1\% & 3.5\%  \\
		\middlehline
		Channel 13  & Bias      & -0.00 &  -0.52 & -0.08 & -0.07 		& -0.14  & -0.29 \\
					& MAE       &  0.80 &   1.17 &  0.84 &  0.83  		&  0.86  &  0.97 \\
					& STD       &  1.00 &   3.04 &  1.13 &  1.10  		&  112  &	1.39 \\
					& Skewness  &  0.02 & -17.85 & -1.14 & -0.98  		& -0.91  & -2.23 \\	
					& Rejection   & -    & -      & -	  & 1.6\%        & 28.1\% & 3.5\%  \\	
		\middlehline
		Channel 15  & Bias      & -0.00 &  -1.28 & -0.07 & -0.06        & -0.33 &  -0.81 \\
					& MAE       &  0.80 &   1.87 &  0.97 &  0.92 		&  1.03 &	1.44 \\
					& STD       &  1.00 &   5.08 &  1.35 &  1.25 		&  1.53 &	2.64 \\
					& Skewness  & -0.00 & -11.15 & -0.60 & -0.83 		& -2.85 &  -4.41 \\ 
					& Rejection & -     & -      & -	 & 5.6\%        & 28.1\% & 3.5\%  \\
		\bottomhline
	\end{tabular}
\end{table*}

@@

The symmetric and low spread error distribution with uncertainty estimates is an important result from the DA perspective. Most of the existing cloud filtering schemes work well only at removing cases with high cloud impact, and as a consequence, the error distributions are highly skewed. To use these observations correctly, data assimilation schemes often inflate their asymmetric error distributions at the cost of artificially suppressing the observational impact. However, the symmetric error distributions obtained from QRNN allow effective utilization more than 94\% of the data without the need for artificial error inflation.
